Spike-and-Slab Generalized Additive Models and Fast Algorithms for High-Dimensional Data

### 2021-06-29 Manuscript/02-Method.Rmd
Revise Method Section

To encourage proper smoothness of the smoothing functions, we adopt the idea of the smoothing penalty from smoothing spline models. The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the frequentist framework. However, direct integration of the smoothness penalty into the sparsity penalty is not obvious. Given that the smoothness penalty matrix is, @Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.

>>>>

To encourage proper smoothness of the smoothing functions, we adopt the idea of the smoothing penalty from smoothing spline models. The basic idea is to set up a smoothing penalty, depending on the smoothness penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density function, or least squared and likelihood function in the penalized framework. However, direct integration of the smoothness penalty into the sparsity penalty is not obvious. Given that the smoothness penalty matrix is most often symmetric and positive semi-definite [^ we limit the discussion to univariate smoothing function. However, the penalty matrix could possibly contain multiple matrices, e.g. tensor product smoother.], the smoothness penalty matrix can be integrated to the design matrix $\bs X$ via re-parameterization. @Marra2011 introduced a re-parameterization of the basis function matrix $\bs X$ based on the decomposition of the smoothness penalty matrix $\bs S$.

### 2021-06-29 Manuscript/SS_GAM.tex
Revise Method Section

To encourage proper smoothness of the smoothing functions, we adopt the
idea of the smoothing penalty from smoothing spline models. The basic
idea is that set up a smoothing penalty, depending on the smoothness
penalty matrix via the penalty introduced in Eq \ref{eq:smoothpen} , in
the posterior density function, or least squared and likelihood function
in the frequentist framework. However, direct integration of the
smoothness penalty with the sparsity penalty is not straightforward.

have the model matrix \(\boldsymbol{X}\) derived from spline functions
of choice \(h_j, j = 1, \dots, p.\) for the \(p\) covariates that are
modeled nonparametrically, the total dimension of \(\boldsymbol{X}\) is
\(n \times \sum\limits_{j=1}^p k_j\). Column-wise centering and
re-parameterization introduced in the GAM as random effect
section{[}\^{}5{]} are performed. The re-parameterization has two
purposes. First of all the re-parameterization reshapes the columns to
be linearly independent to each other, and hence, independent prior can
be set to the coefficients. Secondly, the re-parameterization decomposes
the model matrix \(\boldsymbol{X}_j\) for the \(j\)th covariates to two
sub-matrices representing the null space of the spline and penalized
space of the spline, \[
\boldsymbol{X}_j = \begin{bmatrix}
\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
\end{bmatrix},
\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
denotes the column spaces that need to be shrunk for smoothness of the
spline. The purpose for this re-parameterization is to isolate the null
space from the smoothing space, such that answering the question, if
non-linear relationship is necessary, becomes easy by deciding if the
corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
zero. This strategy is widely used, previously mentioned in
\citet{Scheipl2012} and the references there in.\footnote{Possibly
  select some to add here.} Hereafter, we denote the coefficients of the
null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
for the \(j\)the covariate is \[
f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen},
\] where the superscript \(^t\) denotes matrix transpose.

Without loss of generality, we assume that the model doesn't include the
the parametric part of the model \(\boldsymbol{A}\). The coefficients of
\(\boldsymbol{A}\) can be estimated easily in the proposed EM algorithms
by local update conditioning on the temporary value of the spline
coefficients.

>>>>

To encourage proper smoothness of the smoothing functions, we adopt the
idea of the smoothing penalty from smoothing spline models. The basic
idea is to set up a smoothing penalty, depending on the smoothness
penalty matrix via Equation \ref{eq:smoothpen}, in the posterior density
function, or least squared and likelihood function in the penalized
framework. However, direct integration of the smoothness penalty into
the sparsity penalty is not obvious. Given that the smoothness penalty
matrix is most often symmetric and positive semi-definite {[}\^{} we
limit the discusstion to univariate smoothing function. However, the
penalty matrix could possibly contain multiple matrices, e.g.~tensor
product smoother.{]}, the smoothness penalty matrix can be integrated to
the design matrix \(\boldsymbol{X}\) via re-parameterization.
\citet{Marra2011} introduced a re-parameterization of the basis function
matrix \(\boldsymbol{X}\) based on the decomposition of the smoothness
penalty matrix \(\boldsymbol{S}\).

Column-wise centering and re-parameterization introduced in the GAM as
random effect section{[}\^{}5{]} are performed. The re-parameterization
has two purposes. First of all the re-parameterization reshapes the
columns to be linearly independent to each other, and hence, independent
prior can be set to the coefficients. Secondly, the re-parameterization
decomposes the model matrix \(\boldsymbol{X}_j\) for the \(j\)th
covariates to two sub-matrices representing the null space of the spline
and penalized space of the spline, \[
\boldsymbol{X}_j = \begin{bmatrix}
\boldsymbol{X}_j^{0} & \boldsymbol{X}_j^{pen}
\end{bmatrix},
\] where \(\boldsymbol{X}_j^{null}\) is the column space representing
the null space of the spline model, and \(\boldsymbol{X}_j^{pen}\)
denotes the column spaces that need to be shrunk for smoothness of the
spline. The purpose for this re-parameterization is to isolate the null
space from the smoothing space, such that answering the question, if
non-linear relationship is necessary, becomes easy by deciding if the
corresponding coefficients of the \(\boldsymbol{X}_j^{pen}\) are jointly
zero. This strategy is widely used, previously mentioned in
\citet{Scheipl2012} and the references there in.\footnote{Possibly
  select some to add here.} Hereafter, we denote the coefficients of the
null space as \(\boldsymbol{\beta}^{0}_j\) and the coefficients of the
penalized space as \(\boldsymbol{\beta}^{pen}_j\). The spline function
for the \(j\)the covariate is \[
f_j(X) = f_j^0(X) + f_j^{pen}(X) = {\beta_j^0}^T X_j^0 + {\beta_j^{pen}}^T X_j^{pen}.
\]
